---
title: "Reverse Engineering Project Missing Kids"
author: "Cait Kelley, April Quevedo, Mekhi Abbott, Ilana Williams"
date: "10/5/2023"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [Thousands of kids are missing from school. Where did they go?](https://projects.apnews.com/features/2023/missing-children/index.html)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}

# Clean required data and prepare for analysis if needed. 
library(tidyverse)

# Load janitor
library(janitor)

#load lubridate
library(lubridate)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data

private_enrollment_by_grade <- read_csv("data/private_enrollment_by_grade.csv")
population_by_age <- read_csv("data/population_by_age.csv")
public_enrollment_by_grade <- read_csv("data/public_enrollment_by_grade.csv")

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: “In the 21 states where data was available, [public school] enrollment fell by about 700,000 students between the 2019-2020 and 2021-2022 school years.”

* **Analysis summary**:

Based on the k-12 enrollment totals reported in the 21 states with available data, we were able to confirm enrollment fell by 704593 students between the 2019-2020 school year and the 2021-2022 school year. This matches the exact number listed in the summary file.

```{r}
# Put code to reverse engineer sentence here
states = c("CALIFORNIA","NEW YORK","LOUISIANA","NORTH CAROLINA","WASHINGTON","COLORADO","GEORGIA","MINNESOTA","WISCONSIN","NEW HAMPSHIRE","MASSACHUSETTS","MONTANA","DELAWARE","NORTH DAKOTA","DISTRICT OF COLUMBIA","SOUTH DAKOTA","SOUTH CAROLINA","NEBRASKA","RHODE ISLAND","TENNESSEE","PENNSYLVANIA","FLORIDA") 

public_enrollment_k12 <- public_enrollment_by_grade |>
filter(grade =="k_12_total")|>
  filter(state %in% states)

public_2019_2020 <- public_enrollment_k12 |>
  filter(
    sch_year=="2019-2020"
  )  

public_2021_2022 <- public_enrollment_k12 |>
  filter(sch_year=="2021-2022")
  
colnames(public_2021_2022)[colnames(public_2021_2022) == "enrollment"] <- "enrollment_2021_2022"

colnames(public_2019_2020)[colnames(public_2019_2020) == "enrollment"] <- "enrollment_2019_2020"

public_2019_2020$enrollment_2021_2022 <- public_2021_2022$enrollment_2021_2022

public_2019_2020 <- public_2019_2020 %>%
  select(state, grade, enrollment_2019_2020, enrollment_2021_2022)

public_2019_2020$Difference <- public_2019_2020$enrollment_2021_2022 - public_2019_2020$enrollment_2019_2020

public_enrollment_loss <- sum(public_2019_2020$Difference)
print(public_enrollment_loss)

#Everything we asked ChatGPT:

# 1: Explain in the most simple way how to insert a column into a r studio dataset from another column. Give an example and show the datasets from your example.

# 2: Explain in a simple way how to rename a column in a dataset in r studio

# 3: How do you find the difference between two columns in r

# 4: How to hide columns in a dataset in r studio

# Display results of code below this codeblock

```


### Sentence 2

* **Sentence text**: “Across the states in the analysis, private-school enrollment grew by about 103,000 students.”
* **Analysis summary**: Based on the k-12 enrollment totals reported in the 21 states (plus the District of Columbia) with available data, we were able to confirm enrollment in private schools rose by 102,847 students between the 2019-2020 school year and the 2021-2022 school year. This matches the exact number listed in the summary file.

```{r}
# Put code to reverse engineer sentence here
states = c("CA","NY","LA","NC","WA","CO","GA","MN","WI","NH","MA","MT","DE","ND","DC","SD","SC","NE","RI","TN","PA","FL") 

private_enrollment_k12 <- private_enrollment_by_grade |>
filter(grade =="k_12_total")|>
  filter(state %in% states)

private_2019_2020 <- private_enrollment_k12 |>
  filter(
    sch_year=="2019-2020"
  )  

private_2021_2022 <- private_enrollment_k12 |>
  filter(sch_year=="2021-2022")
  
colnames(private_2021_2022)[colnames(private_2021_2022) == "enrollment"] <- "enrollment_2021_2022"

colnames(private_2019_2020)[colnames(private_2019_2020) == "enrollment"] <- "enrollment_2019_2020"

private_2019_2020$enrollment_2021_2022 <- private_2021_2022$enrollment_2021_2022

private_2019_2020 <- private_2019_2020 %>%
  select(state, grade, enrollment_2019_2020, enrollment_2021_2022)

private_2019_2020$Difference <- private_2019_2020$enrollment_2021_2022 - private_2019_2020$enrollment_2019_2020

private_enrollment_gain <- sum(private_2019_2020$Difference)
print(private_enrollment_gain)

```

### Sentence 3

* **Sentence text**: “The number of children registered for home-schooling surged by around 184,000.”
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

Based on the k-12 enrollment totals reported in the 21 states (including DC) with available data, we were able to confirm homeschool enrollment increased by 184,047 between the 2019-20 school year and the 2021-22 school year. Because the Nebraska data was not available to us, we can assume that homeschool enrollment in NE increased by 5,486. 184,047 + 5,486 = 184,047, matching the number listed in the summary file. 

```{r}
# Put code to reverse engineer sentence here
homeschool_enrollment <- read_csv("data/homeschool_enrollment - Sheet4.csv")
# Display results of code below this codeblock

homeschool_enrollment$Difference <- homeschool_enrollment$enrollment_2021_22 - homeschool_enrollment$enrollment_2019_20

homeschool_enrollment_gain <- sum(homeschool_enrollment$Difference)
print(homeschool_enrollment_gain)
```

### Sentence 4

* **Sentence text**: “The population of school-age kids in the states we studied fell by about 183,000, according to census estimates.”

* **Analysis summary**: We were able to confirm the findings. We found the school-age population fell by 183,486 students in the 21 states + DC which matched the summary file. We followed steps from sentences 1 and 2. We used population data from April 1, 2020 and July 1, 2021. We used the April 1 data because that was prepandemic and excluded data from July 1, 2020.

```{r}
# Put code to reverse engineer sentence here


states = c("CALIFORNIA","NEW YORK","LOUISIANA","NORTH CAROLINA","WASHINGTON","COLORADO","GEORGIA","MINNESOTA","WISCONSIN","NEW HAMPSHIRE","MASSACHUSETTS","MONTANA","DELAWARE","NORTH DAKOTA","DISTRICT OF COLUMBIA","SOUTH DAKOTA","SOUTH CAROLINA","NEBRASKA","RHODE ISLAND","TENNESSEE","PENNSYLVANIA","FLORIDA") 


date_2020_population_by_age <- population_by_age |>
  filter(
    est_date=="2020-04-01")|>
  filter(state %in% states)
    
  

date_2021_population_by_age <- population_by_age|>
  filter(
    est_date=="2021-07-01")|>
  filter(state %in% states)
  

colnames(date_2020_population_by_age)[colnames(date_2020_population_by_age) == "population"] <- "population_2020"

colnames(date_2021_population_by_age)[colnames(date_2021_population_by_age) == "population"] <- "population_2021"

date_2020_population_by_age$population_2021 <- date_2021_population_by_age$population_2021


date_2020_population_by_age$Difference <- date_2020_population_by_age$population_2021 - date_2020_population_by_age$population_2020

date_2020_population_by_age|>
  group_by(state)|>
  summarise(sum=sum(Difference))


population_loss <- sum(date_2020_population_by_age$Difference)
print(population_loss)

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: “That leaves an estimated 230,000 students still unaccounted for in the data.”
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-